---
title: "Seminar02b take home problem"
author: "Olivia"
date: "Monday, January 26, 2015"
output:
  html_document:
    keep_md: yes
---


This solution will be done using the runif() function. 

The first time doing this simulation will be done in the long way. 
I will first define the number of coin flips (sample size) as n and the probablity of the heads coming up as 0.5. 

Simulation will go as below if there are 10 tosses:
```{r}
n1 <- 10
coin1 <- runif(n1, min = 0, max = 1)>0.5
(p1 <- sum(coin1)/n1)

```

This trial indicates that the probability of getting a head in 10 tosses is not exactly 0.5. 

Lets now look at several sample sizes of coin tosses (100, 1000, 10000, 100000, 1000000) and where the probability of getting a head is stil 0.5.
```{r}
n2 <- 100
coin2 <- runif(n2, min = 0, max = 1)>0.5
(p2 <- sum(coin2)/n2)

n3 <- 1000
coin3 <- runif(n3, min = 0, max = 1)>0.5
(p3 <- sum(coin3)/n3)

n4 <- 10000
coin4 <- runif(n4, min = 0, max = 1)>0.5
(p4 <- sum(coin4)/n4)

n5 <- 100000
coin5 <- runif(n5, min = 0, max = 1)>0.5
(p5 <- sum(coin5)/n5)

n6 <- 1000000
coin6 <- runif(n6, min = 0, max = 1)>0.5
(p6 <- sum(coin6)/n6)
```

As we increase the number of coin tosses, it appears that the obtained probability of getting a head is approaching 0.5. 

For ease, the probabilities will be plotted as a scatter-plot. 
```{r}
nsample <- c(n1, n2, n3, n4, n5, n6)
psample <- c(p1, p2, p3, p4, p5, p6)
plot(log10(nsample), psample, type = "p", xlab = 'coin tosses (log10)', ylab = 'probability of getting a head')
abline(0.5,0, col = 'red')

```

As one can see, as the sample size grows, the probability of getting a head approaches 0.5, which is donoted as a red line in the plot above. 

IF I were to change the expected probability of getting a head from 0.5 (ie. a fair coin), to something random like 0.7, I will have to change my code above bit by bit. However, instead of doing that, I would like to look at a higher number of sample sizes with hopefully just a few lines of code. 


```{r}
nsample <- c(10, 20, 40, 60,80,100, 500, 1000, 10000, 100000,1000000)
coinsample <- runif(nsample, min = 0, max = 1)>0.7
psample <-sum(coinsample)/nsample
```
Without showing the probability obtained in each population, as there are so many, I will immediately plot all of my probabilities and sample sizes in a scatter plot. 

```{r}
plot(log10(nsample), psample, type = "p", xlab = 'coin tosses (log10)', ylab = 'probability of getting a head')     
```

The probability of getting a head with an unfair coin will not give a probability of 0.5 as the population gets larger.
